\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[version=4]{mhchem}
\usepackage{stmaryrd}

\begin{document}
\section*{APPLICATION A}
Question : Consider a one-year, fully collateralized commodity futures position on a commodity that experiences a $3 \%$ decline in its spot price. Over that same year, the basis in the commodity fell from the spot price exceeding the futures price by $2 \%$ to a basis of $0 \%$. The riskless interest rate was $4 \%$ and the storage cost of the commodity was $2.4 \%$ per year. Compute the spot return, collateral yield, excess return, roll yield (i.e., roll return), and the return on a fully collateralized position.

\section*{Answer and Explanation}
Here, we are tasked with calculating the different components of return for a commodity futures contract. The variables we need to solve this are here:

\begin{itemize}
  \item Spot Return: this is simply the change in the spot price, or $-3 \%$
  \item Collateral Yield: the collateral yield is equal to the riskless rate, or $4 \%$
  \item Excess Return: this can be broken into the change in the spot price and the change in the basis. The spot price return was $-3 \%$, while the change in basis was equal to $+2 \%$ (the futures price was $2 \%$ lower than the spot price, but now it is equal to it). Therefore, the excess return is $-1 \%$
  \item Roll yield: The roll yield represents the change in basis, which we just solved as $+2 \%$.
\end{itemize}

Note, there are two different ways to calculate the return of the fully collateralized position, which is a good way to check our work:

\begin{itemize}
  \item Equation 1 calculates the return by adding the following: Collateral Yield + Excess Return
  \item Equation 2 expands this equation by adding the following: Spot Return + Collateral Yield + Roll Return
\end{itemize}

Note, excess return can be calculated by adding the Spot Return and Roll Yield.

Either way, we get a 3\% return. Using Equation 1:

Return on Fully Collateralized Position = Collateral Yield + Excess Return $=4 \%+(-1 \%)=3 \%$

Return on Fully Collateralized Position $=$ Spot Return + Collateral Yield + Roll Yield $=(-3 \%)+4 \%+2 \%=3 \%$


\end{document}