\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\begin{document}
\section*{Reading}
Regulatory Environment

Regulation of investments is motivated by concern for the participants directly involved as well as by concern for the overall economy. Privately organized investment vehicles, such as hedge funds, have generally received reduced regulatory scrutiny because the participants involved tend to be sophisticated institutions or individuals perceived to be less in need of regulatory protection than the general public.

Especially since the financial crisis that began in 2007, regulators throughout the world have become increasingly concerned about the role of hedge funds and other investment vehicles in exacerbating systemic risk. Systemic risk is the potential for economy-wide losses attributable to failures or concerns over potential failures in financial markets, financial institutions, or major participants. For example, the collapse of a very large hedge fund may lead to a sequence of collapses and failures that disrupt the financial system and cause widespread economic losses, not so much from the direct asset losses of the collapse as from the inability of the other market participants to trade and manage risks due to the uncertainty that is generated. Regulators are concerned that very large investment funds, such as some hedge funds, or highly complex alternative investment products, such as collateralized debt obligations (CDOs), may increase systemic risk.

\section*{Five Primary Forms of Hedge Fund Regulation}
Regulations of hedge funds take five primary forms:

\begin{enumerate}
  \item Requirements regarding establishing a hedge fund, including registration, licensing, minimum capital, and waiting periods

  \item Registrations or restrictions on investment advisers and hedge fund managers

  \item Restrictions on distribution and marketing of hedge funds, including which marketing channels may be used (e.g., banks), whether advertising is permitted, and to whom funds may be sold

  \item Restrictions on operation of a hedge fund, including leverage, liquidity, risk, reporting, and location of outside service providers

  \item Requirements regarding ongoing reporting

\end{enumerate}

Hedge funds may also be subject to varying levels of taxation and to special taxes, fees, or licensing costs. Understanding regulations is a crucial aspect of alternative investing. Additional information on worldwide regulations is provided in the CAIA Level II curriculum.


\end{document}