\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[version=4]{mhchem}
\usepackage{stmaryrd}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, filecolor=magenta, urlcolor=cyan,}
\urlstyle{same}

\begin{document}
Taxation

Most institutional-quality alternative investments are not created or managed for the primary purpose of avoiding taxes. However, taxation can substantially affect investment returns, and therefore alternative investments are often constructed and managed to prevent additional taxation. In other words, investment pools are formed in light of taxation and with a goal of minimizing the extent to which the pooling of capital increases taxation for the investors relative to direct ownership of the underlying assets. For example, a hedge fund may be domiciled in a particular location for the purpose of preventing additional tax burdens on investors relative to the taxes that would be paid with direct investments using a separately managed and local account. Another hedge fund may be established to invest in municipal bonds for the purpose of generating tax-free income. However, the use of the hedge fund structure and its location do not make the income tax-free. Rather, it is the use of municipal bonds or other tax-free investments, whether inside or outside the hedge fund, that make the income tax-exempt.

In any case, knowledge of general global taxation is helpful in understanding the institutions and other structures involved with alternative investing. The primary objects of taxation throughout the world are income based, wealth based, and transaction based. This section summarizes taxation throughout the world primarily from the perspective of investments.

\section*{Income Taxation}
Throughout the major economies of the world, income is taxed. Income taxation typically includes taxation on individual and corporate income. Most income taxation is progressive. Progressive taxation places higher-percentage taxation on individuals and corporations with higher incomes. Individual income taxation includes taxation of both wage income and investment income.

Although individual wage income and corporate earnings are often fully taxed, the primary issue for investing involves the potential for reduced income tax rates on investment income. Investment income is primarily dividend income, interest income, and capital gains. Investment income from dividends, interest, and capital gains is often either taxed at reduced rates or exempt from income taxation. Although most countries tax all of these types of investment income, the tax rules of individual countries differ primarily by the extent to which dividends, interest, and capital gains are exempted, partially taxed, or fully taxed.

Most major economies, including those of Austria, Brazil, China, Finland, France, Hong Kong, Italy, Japan, the Netherlands, Poland, Sweden, the United Kingdom, and the United States, tax investment income but offer reduced rates on some or all dividends, interest, and capital gains. In the United States, for example, state and municipal bond interest is exempt from federal taxation, and most corporate dividends are taxed at a reduced rate. However, some countries have investment income tax regimes that tax dividends, interest, or capital gains rather heavily or lightly compared to other nations. For example, Canada, Denmark, and Germany tend to have high tax rates on interest income. Australia, Belgium, New Zealand, Switzerland, and Taiwan tend to have low capital gains taxes. ${ }^{1}$ Stephen M. Horan and Thomas R. Robinson, "Taxes and Private Wealth Management in a Global Context," \href{http://www.cfainstitute.org/toolkit}{www.cfainstitute.org/toolkit}, Reading \#70

Other jurisdictions have no income tax or at least no income tax on particular investment pools. These jurisdictions are attractive locations for investment pools in that investors are taxed only by their home country rather than having to pay income taxes on investment income to both their home country and the domicile of the investment fund. These countries include traditional jurisdictions used by hedge funds, such as the Cayman Islands, the British Virgin Islands, Bermuda, Ireland, Luxembourg, Guernsey, and Mauritius. " "Overview of U.S. Asset Management Regulation," by the Regulatory Compliance Association's Senior Fellows from Practice, who are credited as contributing authors to this session.

Some investing offers deferred taxation, in which investment income taxes are not assessed until the funds are withdrawn or distributed. For example, in the United States, qualified retirement savings are generally taxed only at withdrawal. Further, the contributions are often tax-deductible in the period in which the contribution is made. Other opportunities, such as some life insurance contracts, allow tax-deferred accrual of investment income.

Taxation of interest and dividends is generally assessed in the period in which the dividends and interest are distributed. Capital gains tend to be taxed when realized. Capital gains are realized in the period when there is a sale of a security for a price higher than the investor's cost, known as the cost basis. Investments therefore often offer a potentially valuable tax advantage of allowing wealth to be accumulated and accrued through capital appreciation that is not taxed as income until the asset is sold. Further, tax rates may be lower on capital gains, especially when an investment is held for a long time.

Taxation of investment income involves complex rules in most jurisdictions. Understanding taxation can be a very important part of investment management. For example, Section 1256 contracts, which include many futures and options contracts, have potentially enormous tax advantages in the United States, including having their income treated as $60 \%$ long-term capital gain and $40 \%$ short-term capital gain regardless of holding period. Proper decision-making based on this preferential tax treatment can enhance an investor's after-tax return.

\section*{Other Taxes and Withholding}
In most jurisdictions, real estate taxation is an important form of taxation. Often, real estate taxes are assessed by local jurisdictions to fund local services such as schools and governmental services such as law enforcement. The national governments of Australia, Singapore, Belgium, Germany, and the United Kingdom tax real estate. ${ }^{3}$ Horan and Robinson, "Taxes and Private Wealth Management." However, some jurisdictions tax wealth as a general national tax overall. For example, in Colombia, a wealth tax is assessed on all assets, including financial assets. Another important category of taxation is estate taxes that are assessed after the death of a wealthy taxpayer. For wealthy individuals, estate tax rates can be very high.

Although many countries have either drastically reduced or totally eliminated transaction taxes, several European countries continue to impose some form of tax on investment transactions. The United Kingdom uses a stamp tax of $0.5 \%$ on purchases of domestic securities, and France levies a value-added tax on commissions rather than on the transaction value. When market makers trade for their own accounts, they are usually exempted from transaction taxes. In the United States, there is a small fee assessed on securities transactions that is attributed to providing the regulatory services of the SEC.

The international convention on taxing income on foreign investments is to certify that the investor pays taxes to at least one country. Withholding taxes are therefore levied on dividend payments. Although this sometimes results in double taxation, a network of international tax treaties has been signed to prevent double taxation from occurring, so that investors receive a dividend net of withholding tax plus a tax credit from the foreign government but must pay tax on the gross dividends (minus the amount of the withholding tax credit) to the government where they reside. Although this process is potentially lengthy, it allows the investor to\\
reclaim the withholding tax in the foreign country. Depending on the individual country's tax policies, some of the withholding can be retained by the country of origin. Some countries allow tax-free foreign investors (public pension funds) to apply for direct exemptions from tax withholding.


\end{document}