\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[version=4]{mhchem}
\usepackage{stmaryrd}
\usepackage{hyperref}
\hypersetup{colorlinks=true, linkcolor=blue, filecolor=magenta, urlcolor=cyan,}
\urlstyle{same}

\begin{document}
Key Observations Regarding Historical Returns of Funds of Funds

Monthly returns to funds of funds are observed from January of 2000 to December of 2021 for a total of 264 observations. The Statistical Summary of Returns exhibit provides univariate return statistics for six styles of funds of hedge funds.

Statistical Summary of Returns

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\begin{tabular}{l}
Index (Jan. 2000 - Dec. \\
2021) \\
\end{tabular} & \begin{tabular}{l}
HFRI Fund of Funds: \\
Market Defensive Index \\
\end{tabular} & \begin{tabular}{l}
HFRI Fund of Funds: \\
Conservative Index \\
\end{tabular} & \begin{tabular}{l}
HFRI Fund of Funds: \\
Strategic Index \\
\end{tabular} & \begin{tabular}{l}
HFRI Fund of Funds: \\
Diversified Index \\
\end{tabular} & \begin{tabular}{l}
HFRI Fund of Funds: \\
Composite Index \\
\end{tabular} & \begin{tabular}{l}
MSCI World \\
Equity \\
\end{tabular} \\
\hline
\begin{tabular}{l}
Annualized Arithmetic \\
Mean \\
\end{tabular} & $4.1 \%$ & $3.5 \%$ & $4.2 \%$ & $3.9 \%$ & $3.9 \%$ & $6.8 \%$ \\
\hline
\begin{tabular}{l}
Annualized Standard \\
Deviation \\
\end{tabular} & $4.7 \%$ & $3.9 \%$ & $7.2 \%$ & $5.0 \%$ & $5.2 \%$ & $15.4 \%$ \\
\hline
\begin{tabular}{l}
Annualized \\
Semivolatility \\
\end{tabular} & $2.8 \%$ & $4.1 \%$ & $5.9 \%$ & $4.4 \%$ & $4.5 \%$ & $11.8 \%$ \\
\hline
Annualized Median & $3.7 \%$ & $5.2 \%$ & $7.7 \%$ & $5.4 \%$ & $5.4 \%$ & $15.1 \%$ \\
\hline
Skewness & 0.2 & -2.4 & -0.8 & -1.1 & -1.2 & -0.6 \\
\hline
Excess Kurtosis & 0.5 & 12.1 & 3.8 & 4.9 & 5.1 & 1.6 \\
\hline
Sharpe Ratio & 0.3 & 0.3 & 0.3 & 0.3 & 0.3 & 0.3 \\
\hline
Sortino Ratio & 0.6 & 0.6 & 0.6 & 0.6 & 0.6 & 0.4 \\
\hline
\begin{tabular}{l}
Annualized Geometric \\
mean \\
\end{tabular} & $4.0 \%$ & $4.0 \%$ & $4.0 \%$ & $4.0 \%$ & $4.0 \%$ & $5.6 \%$ \\
\hline
\begin{tabular}{l}
First-Order \\
Autocorrelation \\
\end{tabular} & 0.0 & 0.4 & 0.2 & 0.2 & 0.2 & 0.1 \\
\hline
\begin{tabular}{l}
Annualized Standard \\
Deviation \\
(Adjusted for \\
Autocorrelation) \\
\end{tabular} & $4.7 \%$ & $5.7 \%$ & $8.6 \%$ & $6.4 \%$ & $6.6 \%$ & $17.0 \%$ \\
\hline
Maximum & $4.9 \%$ & $2.7 \%$ & $8.8 \%$ & $5.4 \%$ & $5.2 \%$ & $12.8 \%$ \\
\hline
Minimum & $-3.2 \%$ & $-6.8 \%$ & $-9.7 \%$ & $-7.0 \%$ & $-7.6 \%$ & $-19.0 \%$ \\
\hline
Max Drawdown & $-10.9 \%$ & $-20.4 \%$ & $-26.8 \%$ & $-21.8 \%$ & $-22.2 \%$ & $-54.0 \%$ \\
\hline
\end{tabular}
\end{center}

Market-defensive funds of funds tend to have underlying and unhedged short positions. According to HFR, market-defensive funds of funds invest in funds that "generally engage in short-biased strategies such as short selling and managed futures." ${ }^{1}$ ww. \href{http://hedgefundresearch.com/index.php?fuse=indices-str}{hedgefundresearch.com/index.php?fuse=indices-str}. Ideally, this category of funds of funds should have low to negative correlations with respect to major market indices. Ideally, this category of funds of funds should have low to negative correlations with respect to major market indices.

Conservative funds of funds have underlying hedged positions. According to HFR, conservative funds of funds tend to seek consistent returns primarily through "investing in funds that generally engage in more 'conservative' strategies such as Equity Market Neutral, Fixed Income Arbitrage, and Convertible Arbitrage." ${ }^{2}$ Ibid.

Strategic funds of funds tend to have underlying directional bets. According to HFR, strategic funds of funds seek superior returns primarily through "investing in funds that generally engage in more opportunistic strategies such as Emerging Markets, Sector Specific, and Equity Hedge." ${ }^{3}$ Ibid.

Diversified funds of funds represent a broad mix of funds. According to HFR, diversified funds of funds invest "in a variety of strategies among multiple managers." ${ }^{4}$ lbid.

The composite index of funds of funds invests across the various styles of funds of funds investments.

Key observations on funds of funds returns that are consistent with economic reasoning are an essential component of knowledge and include the following:

\begin{enumerate}
  \item All of the fund of funds sub-indices and the composite index exhibited historic return volatility less than one-third that of world equities except the strategic index, which was a little less than one-half.

  \item Only the conservative index exhibited returns that were quite negatively skewed and had very large positive excess kurtosis.

  \item All of the fund of funds sub-indices and the composite index exhibited high first-order return autocorrelation except the defensive index.

  \item All of the fund of funds sub-indices and the composite index exhibited less than half the maximum drawdown of the world equities. The defensive index had a maximum drawdown of only $10.9 \%$, less than one-fifth that of world equities.

\end{enumerate}

In conclusion, we note that returns to funds of hedge funds exhibit volatility and drawdowns substantially lower than those of equity markets. For most styles of funds of hedge funds, autocorrelation is present, showing that many of the underlying assets and strategies may have lower levels of liquidity.


\end{document}